Adapting Duplicate Filters to a Sliding Window Context

### 2020-03-24 qhtv2.tex
add "improvement" graph + fixes in optimal solution

\begin{theorem}
	For $M \geq w (\log_2(w) + 1) \log_2(|\Gamma|)$, the sliding window DDP can be solved exactly (with no errors) in constant time.
\end{theorem}

@@

	The size of the queue is $w\log_2 |\Gamma|$, the size of the dictionary is $w \log_2 |\Gamma| \log_2 w$ (as the dictionary cannot have more than $w$ keys as the same time, and a counter cannot go over $w$, thus being less than $\log_2 w$ bits long). Thus at most $w (\log_2(w) + 1) \log_2(|\Gamma|)$ bits are necessary for this DDF to work.

	Finally this filter does not make any mistake, as the dictionary $D$ keeps an exact tract of how many times each element was present in the sliding window.
\end{proof}
The dependence on $\log_2 |\Gamma|$ can be droped, at the cost of allowing 
errors:
\begin{theorem}
	Let $w \in \mathbb N$. Let $M \simeq 2w\log_2w$, then the sliding window DDP can be solved with almost no error using $M$ bits of memory.

@@

Note that this solution has a time complexity of $O(w)$. Using an additional dictionary, as in the previous proof, we get a filter with an error rate of about $\frac 1w$ and constant time for insertion and lookup, using $5w\log_2 w$ bits of memory.
\end{proof}

@@

\begin{figure}[!ht]
	\input{graphs/compare_queues.tex}
	\caption{Error rate (times 100) of various queuing filters on various sliding window sizes, $M=100000$, $L=10$, $|\Gamma| = 2^{18}$, on a stream of size $10^7$.}\label{tab:comparing_queuing}
\end{figure}
%\begin{table}[]
%	\begin{tabular}{rccccc}
%			   & $w=1000$ & $w=5000$ & $w=10^4$ & $w=10^5$ & $w=10^6$ \\ \hline
%		QHT    & \textbf{9.84} 	  & \textbf{26.25}	 & \textbf{41.09}	 & \textbf{87.89}	  & 70.76      \\
%		Cuckoo & 84.99	  & 99.94	 & 99.95	 & 99.93	  & 99.80      \\
%		SQF    & 16.73 	  & 48.76	 & 69.25	 & 95.77 	  & \textbf{70.14}     \\
%		SBF    & 54.88 	  & 93.93	 & 97.33	 & 97.41	  & 71.17      \\
%	\end{tabular}
%	\caption{Error rate (times 100) of various queuing filters on various sliding window sizes, $M=100000$, $L=10$, $|\Gamma| = 2^{16}$. For each sliding window, the best performance is written in bold.}\label{tab:comparing_queuing}
%\end{table}

@@

\begin{figure}
	\input{graphs/bench_l.tex}
	\caption{Evolution of the error rate of a queueing QHT depending on the $L$ value, for various sliding window sizes, with $M=100000$, $|\Gamma| = 2^{18}$, on a stream of size $10^6$.}\label{fig:bench_l}
\end{figure}

@@

@@

Theoretical results about the queuing construction apply in principe to an infinite stream. However, simulations are necessarily finite, and for very large windows (that are approximately the same size as the whole stream) this causes interesting artifacts in the error rates. Note that these effects have very little impact on practical implementations of queuing filters, since almost all use cases assume a window size much smaller than the stream (or, equivalently, a very large stream). Nevertheless we illustrate the effect of the finite simulation and the parameters affecting it, if only to motivate a further analytical study of this phenomenon.

To accentuate this effect we use the Pearson $\phi$ coefficient that measures statistical correlation (between -1 and +1) between the filter's predictions and the expected answers. A coefficient of $1$ corresponds to an always correct filter, a coefficient of $-1$ to an always incorrect one, and a coefficient of $0$ means the filter produces answers that are statistically independent from the correct ones.

Figure~\ref{fig:strangess} measures $\phi$ as a function of $w$, for different stream sizes $N$. A visible peak in $\phi$ can be found around $w\approx N$.

\begin{figure*}[!ht]
	\centering
	\input{graphs/stream_size.tex}
	\caption{Pearson $\phi$ for queuing SQF ($L = 10$, $M = 10^5$, $|\Gamma| = 2^{16}$) with streams of size $10^5$ to $10^8$.}\label{fig:strangess}
\end{figure*}

>>>>

\begin{theorem}
	For $M \geq w (2\log_2(w) + \log_2(|\Gamma|))$, the sliding window DDP can be solved exactly (with no errors) in constant time.
\end{theorem}

@@

	The size of the queue is $w\log_2 |\Gamma|$, the size of the dictionary is $w (\log_2 |\Gamma| + \log_2 w)$ (as the dictionary cannot have more than $w$ keys at the same time, a dictionary key occupies $\log_2 |\Gamma|$ bits and a counter cannot go over $w$, thus being less than $\log_2 w$ bits long). Thus a requirement of $w (2\log_2(w) + \log_2(|\Gamma|))$ bits for this DDF to work.

	Finally this filter does not make any mistake, as the dictionary $D$ keeps an exact account of how many times each element was present in the sliding window.
\end{proof}
The dependence on $\log_2 |\Gamma|$ can be droped, at the cost of allowing 
errors:
\begin{theorem}
	Let $w \in \mathbb N$. Let $M \simeq 2w\log_2w$, then the sliding window DDP can be solved with almost no error using $M$ bits of memory.

@@

Note that this solution has a time complexity of $O(w)$. Using an additional dictionary, as in the previous proof, but with keys of size $2\log_2(w)$, we get a filter with an error rate of about $\frac 1w$ and constant time for insertion and lookup, using $w \cdot 2\log_2 w + w \cdot (2\log_2(w) + \log_2(w)) = 5w\log_2 w$ bits of memory.
\end{proof}

@@

\begin{figure}[!ht]
	\input{graphs/compare_queues.tex}
	\caption{Error rate (times 100) of various queuing filters on various sliding window sizes, $M=10^5$, $L=10$, $|\Gamma| = 2^{18}$, on a stream of size $10^7$.}\label{tab:comparing_queuing}
\end{figure}
%\begin{table}[]
%	\begin{tabular}{rccccc}
%			   & $w=1000$ & $w=5000$ & $w=10^4$ & $w=10^5$ & $w=10^6$ \\ \hline
%		QHT    & \textbf{9.84} 	  & \textbf{26.25}	 & \textbf{41.09}	 & \textbf{87.89}	  & 70.76      \\
%		Cuckoo & 84.99	  & 99.94	 & 99.95	 & 99.93	  & 99.80      \\
%		SQF    & 16.73 	  & 48.76	 & 69.25	 & 95.77 	  & \textbf{70.14}     \\
%		SBF    & 54.88 	  & 93.93	 & 97.33	 & 97.41	  & 71.17      \\
%	\end{tabular}
%	\caption{Error rate (times 100) of various queuing filters on various sliding window sizes, $M=100000$, $L=10$, $|\Gamma| = 2^{16}$. For each sliding window, the best performance is written in bold.}\label{tab:comparing_queuing}
%\end{table}

@@

\begin{figure}
	\input{graphs/bench_l.tex}
	\caption{Evolution of the error rate of a queueing QHT depending on the $L$ value, for various sliding window sizes, with $M=10^5$, $|\Gamma| = 2^{18}$, on a stream of size $10^6$.}\label{fig:bench_l}
\end{figure}

@@

\begin{figure}
	\input{graphs/improvement.tex}
	\caption{Comparing performances of various filters, in `vanilla' setting or when placed in our queueing structure.}\label{fig:improvement}
\end{figure}

@@

Theoretical results about the queuing construction apply in principle to an infinite stream. However, simulations are necessarily finite, and for very large windows (that are approximately the same size as the whole stream) this causes interesting artefacts in the error rates. Note that these effects have very little impact on practical implementations of queuing filters, since almost all use cases assume a window size much smaller than the stream (or, equivalently, a very large stream). Nevertheless we illustrate the effect of the finite simulation and the parameters affecting it, if only to motivate a further analytical study of this phenomenon.

Figure~\ref{fig:strangess} measures the error rate as a function of $w$, for different stream sizes $N$. A visible decrease in $ER$ can be found around $w\approx N$. While we do not have any explanation for the difference in the peaks sizes and exact location, we give the hypothesis that it is related to the choice of $|\Gamma|$.

\begin{figure*}[!ht]
	\centering
	\input{graphs/stream_size_qht.tex}
	\caption{Error rate for queuing QHT ($L = 10$, $M = 10^5$, $|\Gamma| = 2^{16}$) with streams of size $10^5$ to $10^8$.}\label{fig:strangess}
\end{figure*}
